---
title: "DAPC_Jombart"
author: "Michelle DePrenger-Levin"
date: "November 15, 2016"
output: html_document
---

# add TC
# test 4 population (two of PhFo, PhGi, and PhSc or not)

Packages
```{r}
library(adegenet)
library(pegas)
```

```{r}
rm(list=ls())

```

As opposed to PCA or PCoA/MDS which look at total variance = between groups + within groups, DAPC optomizes between whilc minimizing within    

1. Requires prior groups to be defined. Generally cluster wtih k-means where maximimize between groups variance   
2. So run the clustering k-means algorithm on data transformed through a PCA (instead of raw data) to speed up clustering (fewer variables)    
#Find clusters    
uses genind (microsats) or genlight (SNPs)   
<https://r-forge.r-project.org/scm/viewvc.php/*checkout*/www/files/adegenet-dapc.pdf?revision=899&root=adegenet>   

DAPC -     
    1) different from STRUCTURE, still how proximities of individuals to the different clusters  
    2) Membership probabilities indicate how clear-cut genetic clusters are   
    3) loose clusters = flat distributions of membership probabilities (admixture!)  
    4) Using allele loadings, can predict new indiviudals onto the factorial planes  
    
- transforms data with PCA then performas a discriminant analysis on retained PCs   

infered groups from find.clusters()

Want to retain few PCs without losing infomration (look for asymptote)
 I selected 5000 discriminant functions to retain and it says it saved 23

 customize figure

Compute the contributions of the alleles
```{r}
myPal <- colorRampPalette(c("blue", "gold", "red"))

contrib <- loadingplot(dapc1$var.contr, axis = 2, thres = 0.025, lab.jitter = 1)

```


With troublesome creek  
```{r}
phall <- read.genepop("C:/Users/deprengm/Dropbox/Denver Botanic Gardens Postdoctoral Genetic Files/Phacelia project/Riser_10Oct2016/Genepop/PhfoAllGenePop_13Dec2016.gen")

phall

grp4 <- find.clusters(phall, max.n.clust = 40) #17PCs and 4 groups

```


<http://stackoverflow.com/questions/32704902/discriminant-analysis-of-principal-components-and-how-to-graphically-show-the-di>

Phfo_trunkated_data_75per_Riser the data tab is the structure data

Test Phacelia
```{r}
#Import as Genpop input file
ph1 <- read.genepop("Q:/Research/Lab/Projects/Phacelia formosula/Phfo_GenePop_FinalRISERgenpop.gen")

class(ph1)
ph1

```


#Define useful clusters
```{r}
grp <- find.clusters(ph1, max.n.clust = 35)
# Choose the number of PCs to retain (>=1): 200 - to keep all
# Choose the number of cluster (>=2): 10 - the elbow of the curve

#actual group compared to populations
table(pop(ph1), grp$grp)

table.value(table(pop(ph1), grp$grp), col.lab = 1:10)
```

# How many clusters are useful to describe the data?
```{r}
# #1 transform data using PCA; #2 perform a Discriminant Analysis on retained PCs
dapc1 <- dapc(ph1, grp$grp)
  # 17 - retain as few as possible. STart to gain little more after about 40 PCs
  # 9 - can retain all eigenvalues for small clusters (less than 10s)

dapc1

scatter(dapc1)
```


```{r}
# keep the original 14 populations
dapc(ph1)
dapc.assigned <- dapc(ph1, n.pca = 15, n.da = 13)
scatter(dapc.assigned, scree.da = FALSE, cell = 1.5, 
        bg = "white", cstar=0)

dapc.assigned$var.contr

set.seed(4)

contrib <- loadingplot(dapc.assigned$var.contr, axis = 2,
                       thres = 0.07, lab.jitter = 1)

```

# 14 populations
```{r}
dim(dapc.assigned$posterior)

summary(dapc.assigned) # assign.per.pop lower numbers, not as good assignment to a cluster

compoplot(dapc.assigned, posi = "bottomleft", txt.leg = 1:14,
          ncol = 1, xlab = "individuals")

# Which individuals have less than 90% probability of membership in a single cluster
temp.14 <- which(apply(dapc.assigned$posterior, 1, function(e) all(e < 0.9)))
temp.14

compoplot(dapc.assigned, subset = temp, posi = "bottomright", txt.leg = 1:14, ncol = 2)
```


# 10 groups
```{r}
dim(dapc1$posterior)

summary(dapc1)

compoplot(dapc1, posi = "bottomleft", txt.leg = 1:10,
          ncol = 1, xlab = "individuals")

# Which individuals have less than 90% probability of membership in a single cluster
temp <- which(apply(dapc1$posterior, 1, function(e) all(e < 0.9)))
temp

compoplot(dapc1, subset = temp, posi = "bottomright", txt.leg = 1:10, ncol = 2)
```

#3 groups
```{r}
grp3 <- find.clusters(ph1, max.n.clust = 40)
# 200
# 3

table.value(table(pop(ph1), grp3$grp), col.lab = 1:3)

dapc3 <- dapc(ph1, grp3$grp, n.pca = 15, n.da = 2)
# 15
# 2

scatter(dapc3, scree.da = FALSE, cell = 1.5, bg = "white", cstar =0)

summary(dapc3)

```


# Two groups
```{r}
dim(dapc2$posterior)

summary(dapc2)

compoplot(dapc2, posi = "bottomleft", txt.leg = 1:2,
          ncol = 1, xlab = "individuals")

# Which individuals have less than 90% probability of membership in a single cluster
temp.2 <- which(apply(dapc2$posterior, 1, function(e) all(e < 0.9)))
temp.2

compoplot(dapc2, subset = temp.2, posi = "bottomright", txt.leg = 1:2, ncol = 2)

```




```{r}
#Minimum spanning tree on squared distances between populations in entire space
myCol <- c("darkblue", "purple", "green", "orange", "red", "blue",
           "brown","darkgreen","deepskyblue4","cornsilk4") #,"deeppink4","gray30","coral2")


scatter(dapc1, ratio.pca = 0.3, bg = "white", pch = 20, cell = 0,
        cstar = 0, col = myCol, solid = 0.4, cex = 3, clab = 0, mstree = TRUE,
        scree.da = FALSE, posi.pca = "bottomright", leg = TRUE, 
        txt.leg = paste("Cluster", 1:10))
par(xpd = TRUE)
points(dapc1$grp.coord[, 1], dapc1$grp.coord[, 2], pch = 4, cex = 2,
       lwd = 8, col = "black")
points(dapc1$grp.coord[, 1], dapc1$grp.coord[, 2], pch = 4, cex = 2,
       lwd = 2, col = myCol)
```

```{r}
#single discriminant function like when K=2
scatter(dapc1, 1,1, col = myCol, bg= "white", scree.da = FALSE, legend = TRUE,
        solid = 0.4)


```

#Contribution of alleles
```{r}



```















#Assume can be split into two groups


```{r}
grp <- find.clusters(ph1, max.n.clust = 40)
# Choose the number of PCs to retain (>=1): 200 - to keep all
#Choose the number of cluster (>=2): 13 - the elbow of the curve
names(grp)

grp$size

#actual groups
table(pop(ph1), grp$grp)

table.value(table(pop(ph1), grp$grp), col.lab = 1:13)
```

# Are two many clusters are useful to describe the data?
```{r}

grp2 <- find.clusters(ph1, max.n.clust = 40)
# Choose the number of PCs to retain (>=1): 200 - to keep all
#Choose the number of cluster (>=2): 2 - the elbow of the curve (10 might be best)
names(grp2)

grp2$size

#actual groups
table(pop(ph1), grp2$grp)

table.value(table(pop(ph1), grp2$grp), col.lab = 1:2)

# #1 transform data using PCA; #2 perform a Discriminant Analysis on retained PCs
dapc2 <- dapc(ph1, grp2$grp)
  # 10 - retain as few as possible. STart to gain little more after about 17 PCs
  # 1 - can retain all eigenvalues for small clusters (less than 10s)

dapc2 # describes way less of the variance - proportion of conserved variance

scatter(dapc2)
```


#Locations of individuals
```{r}


gpsdata <- read.csv("Q:/Research/Lab/Projects/Phacelia formosula/Project Information/Phacelia project information with GPS data 2016_medl.csv")

head(gpsdata)

length(ph1@pop) #395
length(gpsdata$Code) #421
table(ph1@pop)
table(gpsdata$Code)

#Set level order the same
levs <- unique(unlist(lapply(ph1@pop, levels)))
levels(gpsdata$Code) <- levs



```





retain all PCs when k-means clustering   
after k groups, do DAPC but don't retain more than 1/3 of the total number of PCs, retain all of the DFs (what's a DF?)   
optim.a.score() to determine best number of PCs to retain

