---
title: "IBD"
author: "Michelle DePrenger-Levin"
date: "November 17, 2016"
output: html_document
---

```{r}
library(raster)
library(rgdal)
library(maptools)
library(devtools)
library(ecodist)
library(vegan)
library(prism)
library(dismo)


```


ape
ade4

Isolation by distance (IBD) and Isolation by environment (IBE)
1) Mantel test
vegan - Forester et al 2016    
spatial eigenfunction analysis and multivariate linear regression (Diniz-Filho et al. 2013 and Legendre et al. 2015)  
i - PCoA to pairwise multivariate genetic distance matrix (Bray-Curtis)
ii - keep PCoA axes based on the broken-stick criterion (Legendre & Legendre 2012)   
iii - response data = kept axes, for redundancy analysis - mutivariate linear regression (RDA)
iv - predictors are spatial eigenfunctions (distance-based Moran's eigenvector maps, dbMEMs)
v - forward selection to reduce number of dbMEMs (Blanchet et al. 2008)
vi - calcuate r^2 assess signficantion using 1000 permutations   
<https://github.com/rforge/sedar/tree/master/pkg> 
both PCNM and packfor
```{r}
library(devtools)
#install.packages("vegan")
#install_github("vegandevs/vegan")
library(vegan)
#install.packages("PCNM") # not avaliable for 3.3.1, a funciton in vegan?
?pcnm
library(boot)
?packfor
#install_git("rforge/sedar/pkg/AEM")
#install_github("rforge/sedar/pkg/PCNM")

#install.packages("ecodist")
library(ecodist)



```


soils <https://www.nrcs.usda.gov/wps/portal/nrcs/detail/soils/survey/?cid=nrcs142p2_053627>
Data from USDA <http://websoilsurvey.nrcs.usda.gov/app/WebSoilSurvey.aspx>   
 1. Analsis Tools -> Overlay -> Spatial Join 2. In spatial join dialog box: -Target features = Points -Join features = Polygons -Join Operation = Join one to one -Right click on 'yield' field in join feature field map, select properties, choose 'Maximum' as merge rule -Match option = Intersect 
Try to get netCDF data for these points
```{r}
soils <- read.delim("P:/hackathon/Phacelia/Walden_Phacelia/SoilsPerPoints.txt",
                    sep = ",")

head(soils)
soils$Population <- gsub('[[:digit:]]+', '', soils$Comment)
soils$Population[soils$Comment == " "] <- "TroublesomeCreek"
unique(soils$Population)

soils$Population[grep("WSE", soils$Population)] <- "WeatherStation"
soils$Population[grep("Brownlee", soils$Population)] <- "Brownlee"
soils$Population[grep("RNA", soils$Population)] <- "ResearchNaturalArea"
soils$Population[grep("iamond", soils$Population)] <- "DiamondJNE"

```


Diniz-Filho et al 2013   
```{r}
library(raster)
library(prism)

options(prism.path = "Q:/Research/All_Projects_by_Species/Astragalus SPECIES/Astragalus_microcymbus/PRISM_asmiclimate")

#Precipitation total, rain and snow
get_prism_monthlys(type="ppt", mon = 1:12, keepZip = FALSE, years = 1980:1993)

#Maximum temperature (average)
get_prism_monthlys(type="tmax", mon = 1:12, keepZip = FALSE, years = 1980:1993)

#Minimum temperature (average)
get_prism_monthlys(type="tmin", mon = 1:12, keepZip = FALSE, years = 1980:1993)

ph.ll <- spTransform(SpatialPoints(cbind(soils$EastingAll,soils$NorthingAll),
                                   proj4string = CRS("+proj=utm +zone=13 + datum=WGS84")),
                     CRS("+proj=longlat +datum=WGS84"))
```

```{r}

ph.cols <- data.frame(coordinates(ph.ll),soils$Population,soils$Ind_Num)


maxs <- grep("tmax", ls_prism_data(absPath=TRUE)[,2])
mins <- grep("tmin", ls_prism_data(absPath = TRUE)[,2])
ppt <- grep("ppt", ls_prism_data(absPath = TRUE)[,2])

avgTemps.max <- lapply(maxs, function(x){
  rastertemps <- raster(ls_prism_data(absPath=TRUE)[x,2])
  data.frame(data = extract(rastertemps, ph.cols[,c("coords.x1","coords.x2")]), 
             date = ls_prism_data()[x,])
})

avgTemps.min <- lapply(mins, function(x){
  rastertemps <- raster(ls_prism_data(absPath=TRUE)[x,2])
  data.frame(data = extract(rastertemps, ph.cols[,c("coords.x1","coords.x2")]), 
             date = ls_prism_data()[x,])
})

avgPrecip <- lapply(ppt, function(x){
  rastertemps <- raster(ls_prism_data(absPath=TRUE)[x,2])
  data.frame(data = extract(rastertemps, ph.cols[,c("coords.x1","coords.x2")]), 
             date = ls_prism_data()[x,])
})

mins.avg <- do.call(rbind, avgTemps.min)
maxs.avg <- do.call(rbind, avgTemps.max)
ppt.avg <- do.call(rbind, avgPrecip)


```

<https://cran.r-project.org/web/packages/dismo/vignettes/sdm.pdf>   
The code above returns a raster with the 18 bioclimate variables covering the whole world with a resoltion of 2.5 minutes of degrees:

BIO1 = Annual Mean Temperature
BIO2 = Mean Diurnal Range (Mean of monthly (max temp – min temp))
BIO3 = Isothermality (BIO2/BIO7) (* 100)
BIO4 = Temperature Seasonality (standard deviation *100)
BIO5 = Max Temperature of Warmest Month
BIO6 = Min Temperature of Coldest Month
BIO7 = Temperature Annual Range (BIO5-BIO6)
BIO8 = Mean Temperature of Wettest Quarter
BIO9 = Mean Temperature of Driest Quarter
BIO10 = Mean Temperature of Warmest Quarter
BIO11 = Mean Temperature of Coldest Quarter
BIO12 = Annual Precipitation
BIO13 = Precipitation of Wettest Month
BIO14 = Precipitation of Driest Month
BIO15 = Precipitation Seasonality (Coefficient of Variation)
BIO16 = Precipitation of Wettest Quarter
BIO17 = Precipitation of Driest Quarter
BIO18 = Precipitation of Warmest Quarter
BIO19 = Precipitation of Coldest Quarter
```{r}
install.packages(c('raster', 'rgdal', 'dismo', 'rJava'))

library(dismo)
#check out example file
file <- paste(system.file(package="dismo"), "/ex/bradypus.csv", sep="")
bradypus <- read.table(file, header=TRUE, sep=",")
head(bradypus)

phal <- gbif("phacelia", "alba*", geo=FALSE)
phde <- gbif("phacelia", "denticulata*", geo=FALSE)
phha <- gbif("phacelia", "hastata*", geo=FALSE)
phhe <- gbif("phacelia", "heterophylla*", geo=FALSE)
phse <- gbif("phacelia", "sericea*", geo=FALSE)

phgeo <- subset(phse, !is.na(phse$lon) & !is.na(phse$lat))


install.packages("maptools")
library(maptools)

data(wrld_simpl)
plot(wrld_simpl, xlim=c(-150,-100), ylim=c(20,70), axes=TRUE,
     col="light yellow")
points(phgeo$lon, phgeo$lat, col="orange", pch=20, cex=0.75)


bioclim <- getData("worldclim", var = 'bio', res = 2.5)

    
usa <- getData("GADM", country="USA", level=2)
colorado <- usa[usa$NAME_1 == "Colorado",]
counties <- c("Larimer","Jackson","Grand")
ph.counties <- colorado[colorado$NAME_2 %in% grep(paste(counties, collapse="|"),
                                                  colorado$NAME_2, value=TRUE),]
ph.counties <- ph.counties[ph.counties$NAME_2 != "Rio Grande",]
plot(usa)
plot(bioclim$bio1, main="Annual Mean Temperature")

#SRTM 90 Elevation
srtm <- getData('SRTM', lon= -106, lat = 41)
#mosaic enough tiles to gt them together!
srtm1 <- getData('SRTM', lon= -106, lat = 40)
strmmosaic <- mosaic(srtm, srtm1, fun=mean)

plot(strmmosaic)
plot(ph.counties, add=TRUE)

ph.p <-data.frame(coordinates(ph.points))


```


Data cleaning
```{r}
sum(duplicated(soils[,c("EastingAll","NorthingAll")]))
#28

soils.unique <- soils[!duplicated(soils[,c("EastingAll","NorthingAll")]),]
ph.points <- spTransform(SpatialPoints(cbind(soils.unique$EastingAll,soils.unique$NorthingAll),
                                       proj4string = CRS("+proj=utm +zone=13 +datum=WGS84")),
                         CRS("+proj=longlat +datum=WGS84"))

#cross check coordinates
#Make a SpatialPointsDataFrame
library(sp)


ggplot()+
  geom_path(data=ph.counties, aes(x=long, y=lat, group=group))+
  coord_map()+
  geom_point(data=ph.p, aes(x=coords.x1, y=coords.x2))


```

Sampling bias    
with many samples, subsampling to reduce sampling bias   
ph.points == acg in example
```{r}
r <- raster(ph.points)
#cells are set to 1 degree
res(r) <- 1
# extend the extent of the RasterLayer a little
r <- extend(r, extent(r)+1)
#sample
phsel <- gridSample(ph.points, r, n=1) # must be 1 per degree cell

p <- rasterToPolygons(r)
plot(p, border='gray')
points(ph.points)
```

Background data - for when you don't have absence data   
OR pseudo-absences but these guys don't like as much
dismo to sample random points for background data
```{r}

files <- list.files(path=paste(system.file(package="dismo"), '/ex',
                               sep=''), pattern='grd', full.names=TRUE )

#mask to exclude areas with no data
mask <- raster(files[1])

#select random sample points, set seed
set.seed(1963)
bg <- randomPoints(mask, 500)

plot(!is.na(mask), legend=FALSE)
points(bg, cex=0.5)

#extent to restrict areas from which random points will be drawn
e <- extent(-80, -53, -39, -22)
bg2 <- randomPoints(mask, 50, ext=e)
plot(!is.na(mask), legend=FALSE)
plot(e, add=TRUE, col='red')
points(bg2, cex=0.5)


#Phacelia
rm(mask,bg2,rr)
mask <- raster(xmn=-107, xmx=-105,ymn=40, ymx = 41, nrow=100, ncol=100)
mask[]=runif(100*100)

plot(mask)
plot(ph.counties, add=TRUE)
points(ph.points)

rr <- mask(mask,ph.counties)

plot(rr)
plot(ph.counties, add=TRUE)

bg2 <- randomPoints(rr, 500)
plot(!is.na(mask), legend=FALSE)
plot(rr, add=TRUE, col='red')
points(bg2, cex=0.5)

#circles with a radius of 5 km
x <- circles(rasterToPoints(rr), d=5000, lonlat=TRUE)
pol <- polygons(x)

head(x@presence)
head(x@absence)

# then use rgeos package to dissolve circles (remove boundaries where overlap)
#sample from circles
samp1 <- spsample(pol, 250, type="random", iter=25)
cells <- cellFromXY(mask, samp1)
cells <- unique(cells)

xy <- xyFromCell(mask, cells)

plot(pol, axes=TRUE)
points(xy, cex=0.75, pch=20, col='blue')
```

Raster data - predictor variables    
ASCII files are slow
```{r}
files <- list.files(path=paste("Q:/Research/All_Projects_by_Species/aa_Shapefiles_Maps/aa_GENERAL_non-species_files/All_General_Background_Layers/World BioClim Date/bio_2-5m_bil", 
                               sep = ''),
                    pattern='grd', full.names=TRUE)

# Don't know what the .grd.vat.cpg files are
predictors <- stack(files[c(1,4:21)])
#missing projection
proj4string(predictors) <- CRS("+proj=longlat +datum=WGS84")

names(predictors)
plot(predictors)

predictors <- crop(predictors, extent(r)+1)

plot(predictors)

plot(predictors,1)
plot(ph.counties, add=TRUE)
points(ph.points, col="blue", pch =20, cex = 0.5)

```

Create new predictor variables - example: use land cover data and focal() and create new variable that indicates how much forest area is available within Xkm of a grid cell..
```{r}
?focal

```

#Extract values from rasters
```{r}
presvals <- extract(predictors, ph.points)

#set random seed to always create same random set of points
set.seed(0)

backgr <- randomPoints(predictors, 500)

absvals <- extract(predictors, backgr)

# row for each point 
pb <- c(rep(1, nrow(presvals)), rep(0, nrow(absvals)))
site <- c(soils.unique$Population, rep(0, nrow(absvals)))
sdmdata <- data.frame(cbind(site, pb, rbind(presvals, absvals)))

#categorical variables   
#SKIP
sdmdata[,'biome'] <- as.factor(sdmdata[,'biome'])

head(sdmdata)

pairs(sdmdata[,2:20], cex=0.1, fig=TRUE)

#too many, take out highly correlated ones
cor.table <- cor(sdmdata[,2:20])
```






<https://www.rforge.net/doc/packages/climates/html/bioclim.html>
```{r}
install.packages("climates") #nope, how?!


```





follow vegan tutorial, trying to get environmental variables   
<http://cc.oulu.fi/~jarioksa/opetus/metodi/vegantutor.pdf>


redundancy analysis (RDA): combines PCA on allele frequencies and multiple regression
Meirmans Molecular Ecology "Seven common mistakes in pouplaiton genetics and how to avoid them"