---
title: "Morphometrics"
author: "Michelle DePrenger-Levin"
date: "November 11, 2016"
output: pdf_document
---

```{r, eval=FALSE}
#install.packages("shapes")
library(shapes)
library(MASS)
library(cluster)
library(dendextend)
library("ape")
library(ggplot2)
library(vegan)
```


<http://life.bio.sunysb.edu/morph/>    
> software    
  > Utility programs    
  Download tpsUtil  
  > Thin-plate spline
  Downlaod tpsRelw   
  Download tpsSmall     

<http://swordtail.tamu.edu/anyfish/Modeling_body_and_fin_shape_%28morphometrics%29>
1. Build a TPS file using tpsUtil   
2. Open tpsUtil    
  * Build tps file from images      
  * Input directory -> go to file with images   
  * Output to create file where you'll save info  
  * Setup - check all of images you want, create   
  
3. Can view that setup file in a text editor   
  * LM=0 : "landmark ID"    
  * info for each image selected     
  * after the landmarks are set, that LM will have numbers for each.    
  
4. Landmark tps files    
  * Open tpsDig    
  * Open tps file    
  * Input source, open the tps file   
  * Set scale: Options>Image tools or tool bar picture with tools images    
  ** a) Measure tab    
        Reference length - scale to image   
        Set scale > click and click along ruler   
     b) Apply landmark > Digitize landmarks (circle and cross image button)    
        cursor becomes circle cross   
        click on landmarks - the first and 22nd are identical (or some number - the midway?)    
        match landmarking guide - exact order of landmarks in each image    
        number and order must match all     
        File > Save data > save and overwrite  
      c) Visulaization aids    
        * Menu > preliminatries > option: create or edit wireframe   
        * Menu > file > option: import outline file    
      d) Convert TPS into NTS     
        * TPSUtil, convert TPS/NTS file   
        * check box for using the sacale factor    
        * check box for using image names as labels  
      e) Procrustes superimposed   
        * Menu > preliminatries> option: Procrustes fit  
        
5. TPSSmall - test whether variation in shape is too large   
  * regresses through the origin the set of Euclidean distances in the Euclidean space onto the set of Procrutes shape distances      
  * want approximation to give regression with both slope and correlation virtually equal to 1    

6. Consensus file (An average of a population)   
  * Open tpsRelw    
      * Landmark all LMs, combine and generate a consensus for that population   
  * Data > open the TPS file   
  * Compute > Consensus   
    * Display > Consensus   
      * see the cons   
      * File > Save > Save consensus... > Save with ".TPS"    


7. MorphoJ
  * read NTS files in with NTSYSpc format and with symmetry since both sides were digitized
  * Do procrustes fit - to project into tangent space because standard statistical methods such as regression, AOV require data in flat Euclidean space. The point of tangency between two spaces is the sample mean shape. 
  
  Check with TPSSmall a regression through origin the set of Euclidean distances in the eudlidean space onto the set of Proscrustes shape distances - if approximation is good, regression slope and correlation will be nearly equal to 1. 
  <http://www.flywings.org.uk/MorphoJ_guide/frameset.htm?index.htm> 
  
  USE ANOVA with species?? as main effect, populaiton and leaves as random effects for a Procrustes ANOVA to check for variation in with and among populaitons and whither this is more than explained by measuremetn error. - must be hierarchical ie. taking into account the nested structure of data since we have unbalanced design In MorphoJ: 'variation' and 'Procrustes ANOVA'
  - look at sum of squares to measure deviation of observations from the means (within groups)   
  
  Permutation tests and discriminant analyses: taking the Procrustes ANOVA - if lowest level, leaves, is statisctically signficant and larger than measurement error. then if more variation in leaves of different populations than within the same population - then pool and use average per population.  
  1. Parametric t-test for independent samples   
  2. repeat test using permutations (which do not assume normally distributed data) Useing PAST with Statistics and F and T test for two samples
  3. Permutation tests for group differences in MorphoJ using regression approach a) preliminaries, edit covariates and code one pop '1' and one '-1'. Then size is regressed onto dummy covariate using permutations to test significance. get P value and percent variance explained by population. 
  
  
  
  

Landmark data
```{r}
sites <- c("AS","BR","FR","RN","TC")

landmarks <- lapply(sites, function(x){
  site <- read.csv(paste("Q:/Research/All_Projects_by_Species/Phacelia SPECIES/Phacelia_formosula/Phacelia formosula_abiotic/Landmark_",x,".csv", sep=""), header=FALSE)
  sitexy <- site[complete.cases(site),]
  colnames(sitexy) <- c("X","Y")
  ind <- seq(23,nrow(site),25)
  indivduals <- sapply(gsub(".*[=]|[.].*", "", site[ind,1]), function(x) rep(x,21))
  sitexy$Ind <- c(indivduals)
  sitexy$Ldm <- rep(1:21,nrow(sitexy)/21)
  sitexy
})

# how many individuals per site
indPsite <- lapply(landmarks, function(x){
  length(unique(x$Ind))
})

ldms <- do.call(rbind,landmarks)
head(ldms)

# output each population landmarks
wide.AS <- reshape(landmarks[[1]], idvar="Ind", timevar = "Ldm", direction = "wide") 
wide.BR <- reshape(landmarks[[2]], idvar="Ind", timevar = "Ldm", direction = "wide") 
wide.FC <- reshape(landmarks[[3]], idvar="Ind", timevar = "Ldm", direction = "wide") 
wide.RN <- reshape(landmarks[[4]], idvar="Ind", timevar = "Ldm", direction = "wide") 
wide.TC <- reshape(landmarks[[5]], idvar="Ind", timevar = "Ldm", direction = "wide") 

write.table(wide.AS[,-1], "C:/Users/deprengm/Dropbox/Denver Botanic Gardens Postdoctoral Genetic Files/Phacelia project/Phacelia 2014 collection and morphology data/wide_AS.txt",
            col.names=FALSE,row.names = FALSE, quote=FALSE)
write.table(wide.BR[,-1], "C:/Users/deprengm/Dropbox/Denver Botanic Gardens Postdoctoral Genetic Files/Phacelia project/Phacelia 2014 collection and morphology data/wide_BR.txt",
            col.names=FALSE,row.names = FALSE, quote=FALSE)
write.table(wide.FC[,-1], "C:/Users/deprengm/Dropbox/Denver Botanic Gardens Postdoctoral Genetic Files/Phacelia project/Phacelia 2014 collection and morphology data/wide_FC.txt",
            col.names=FALSE,row.names = FALSE, quote=FALSE)
write.table(wide.RN[,-1], "C:/Users/deprengm/Dropbox/Denver Botanic Gardens Postdoctoral Genetic Files/Phacelia project/Phacelia 2014 collection and morphology data/wide_RN.txt",
            col.names=FALSE,row.names = FALSE, quote=FALSE)
write.table(wide.TC[,-1], "C:/Users/deprengm/Dropbox/Denver Botanic Gardens Postdoctoral Genetic Files/Phacelia project/Phacelia 2014 collection and morphology data/wide_TC.txt",
            col.names=FALSE,row.names = FALSE, quote=FALSE)


#each column as a landmark
head(ldms)

wide.ldms <- reshape(ldms, idvar="Ind", timevar = "Ldm", direction = "wide")

head(wide.ldms)
write.table(wide.ldms[,-1], "C:/Users/deprengm/Dropbox/Denver Botanic Gardens Postdoctoral Genetic Files/Phacelia project/Phacelia 2014 collection and morphology data/Phfo_ldms.txt", col.names = FALSE, row.names = FALSE, quote=FALSE)
```


#Shapes
<http://www.indiana.edu/~g562/PBDB2013/Day%202B%20-%20Geometric%20Morphometrics%20in%20R.pdf>
<https://www.maths.nottingham.ac.uk/personal/ild/shapes/quick-tutorial.txt> 
```{r}
k <- 21 #number of points
m <- 2  #number of dimensions
new.dat <- read.in("C:/Users/deprengm/Dropbox/Denver Botanic Gardens Postdoctoral Genetic Files/Phacelia project/Phacelia 2014 collection and morphology data/Phfo_ldms.txt",k,m)

plotshapes(new.dat)
out.procf <- procGPA(new.dat)
plotshapes(out.procf$rotated)

shapepca(out.procf) #plots the first few PCs

# by population
new.dat.list <- lapply(c("wide_AS","wide_BR","wide_FC","wide_RN","wide_TC"), function(x){
  read.in(paste("C:/Users/deprengm/Dropbox/Denver Botanic Gardens Postdoctoral Genetic Files/Phacelia project/Phacelia 2014 collection and morphology data/",x,".txt", sep=""),k,m)
})

out.procf.site <- lapply(new.dat.list, function(x) {
  procGPA(x)
  })

out.procf.site[[1]]$n

par(mfrow=c(3,2))
par(mar=c(2,2,2,2))
lapply(out.procf.site, function(x){
  plotshapes(x$rotated)
})



# the Riemannian shape distance between the mean shapes
sites
comparisons <- t(combn(1:5,2))
comparisons[10,][2]
reimannian.site <- lapply(1:10, function(x){
  r.s <- riemdist(out.procf.site[[ comparisons[x,][1] ]]$mshape, 
                  out.procf.site[[ comparisons[x,][2] ]]$mshape)
  out <- cbind(paste(sites[ comparisons[x,][1] ],sites[ comparisons[x,][2] ], sep="-"),
               r.s)
  })
r.s <- do.call(rbind,reimannian.site)
data.frame(r.s)

## The procustes distance
proc.site <- lapply(1:10, function(x){
  p.s <- procdist(out.procf.site[[ comparisons[x,][1] ]]$mshape, 
                  out.procf.site[[ comparisons[x,][2] ]]$mshape)
  out <- cbind(paste(sites[ comparisons[x,][1] ],sites[ comparisons[x,][2] ], sep="-"),
               p.s)
})

p.s <- do.call(rbind, proc.site)
p.s

# tests for 2D mean shape difference, including boostrap and permutation tests
testshapediff <- lapply(1:10, function(x){
  resampletest(new.dat.list[[ comparisons[x,][1] ]],
               new.dat.list[[ comparisons[x,][2] ]]) #resamples = 200, bootstrap with replacement
  })

head(testshapediff[[1]])

foo <- testshapediff
  
foo1 <- lapply(1:10, function(x){
  out <- paste(sites[ comparisons[x,][1] ],sites[ comparisons[x,][2] ], sep="-")
  list(testshapediff[[x]],out)
  })

foo1[[1]][[2]][1]

do.call(rbind,lapply(foo1, function(f){
  data.frame(SiteComp = f[[2]][1],
             Lambda = round(f[[1]]$lambda,2),
             Lambda.pvlaue = round(f[[1]]$lambda.pvalue,2),
             James = round(f[[1]]$J,2),
             James.pvalue = round(f[[1]]$J.pvalue,2))
  })
)



```





### Include TC
```{r}
phdaisy_wide <- daisy(wide.ldms[,-1], metric="gower")
phclust_wide <- hclust(phdaisy_wide)

ph.dendwide <- as.dendrogram(phclust_wide)

plot(phclust_wide, labels = wide.ldms$Ind, cex = 0.75)

par(mar=c(5,5,2,2))
ord <- isoMDS(phdaisy_wide)
code <- substr(wide.ldms$Ind,1,2)

ord.xy <- data.frame(wide.ldms$Ind,ord$points,code)
colnames(ord.xy) <- c("Individual","X","Y","Site")

ggplot(ord.xy, aes(X,Y,colour=Site))+
  geom_point()+
  stat_ellipse()

```


### Exclude TC
# plant height and width (Atwood?) 
# go back and measure seed entire or pitted FC and TC and detect differences? add to morph data
#abiotic envelope

# Morphometic and DAPC including TC and without
biological conservation
diveristy and distributions - conservation and testing a species hypothosis, what lines of evidence needed to split vs. lump... 

Edwards Knowles 2014 PRSB

# First keep leaf landmarks separate from landmarks
Do not include plant height or width
```{r}
phfo <- read.csv("Q:/Research/Lab/Projects/Phacelia formosula/Morphology Images_phfo/Phacelia.formosula.data.2014CSV.csv")

head(phfo)
phfo <- phfo[,-(18:19)]

include <- names(phfo)[c(10:17)]

pairs(phfo[,include])
cor(phfo[,include], method = "pearson", use = "na.or.complete")

ph.daisy.5 <- daisy(phfo[,include[-1]], metric = "gower") #removed leaf length, PCC 0.62 with leaf width
phclust.all.5 <- hclust(ph.daisy.5)

ord.5 <- isoMDS(ph.daisy.5)
ord.xy5 <- data.frame(phfo,ord.5$points)

ggplot(ord.xy5, aes(X1,X2,colour=Population.Code))+
  geom_point()+
  stat_ellipse()+
  theme_bw()

TukeyHSD(aov(ord.xy5$X1 ~ ord.xy5$Population.Code))


######## Without TC
pairs(phfo[-grep("TC",phfo$Population.Code),include[-1]])

ph.daisy.4 <- daisy(phfo[-grep("TC",phfo$Population.Code),include[-1]], metric = "gower") 
  #removed leaf length, PCC 0.62 with leaf width
phclust.all.4 <- hclust(ph.daisy.4)

ord.4 <- isoMDS(ph.daisy.4)
ord.xy4 <- data.frame(phfo[-grep("TC",phfo$Population.Code),],ord.4$points)

ggplot(ord.xy4, aes(X1,X2,colour=Population.Code))+
  geom_point()+
  stat_ellipse()+
  theme_bw()

TukeyHSD(aov(ord.xy4$X1 ~ ord.xy4$Population.Code))


##########
ph.dist.4 <- dist(phfo[-grep("TC",phfo$Population.Code),include[-1]])
ph.dist.5 <- dist(phfo[,include[-1]])

meta.5 <- metaMDS(comm = ph.daisy.5) #no solution reached for 4
meta.4 <- metaMDS(comm = ph.dist.4) #solution reached

plot(meta.5)
plot(meta.4)
```


Combined   
Merge morphometrics 
```{r}
# Change population code to match

unique(phfo$Population.Code)
unique(code)
wide.ldms$Ind <- sapply(wide.ldms$Ind, function(x) gsub("FR","FC",x))

ph.merg <- merge(wide.ldms, phfo, by.x = "Ind", by.y = "Individual.ID")
names(ph.merg)

table(ph.merg$Population.Code)

names(ph.merg)
names(ph.merg[,c(2:43,50:59)])

names(ph.merg[,c(2:43,52:59)])

ph.daisy <- daisy(ph.merg[,c(2:43,52:59)], metric = "gower") # take out 50 and 51 = plant height and width, and leaf length for being correlated
phclust.all <- hclust(ph.daisy)

plot(phclust.all, labels = ph.merg$Species)
plot(phclust.all, labels = ph.merg$Population.Code)


ord.merged <- isoMDS(ph.daisy)
ord.merged <- data.frame(ph.merg, ord.merged$points)

ggplot(ord.merged, aes(X1,X2,colour=Population.Code))+
  geom_point()+
  stat_ellipse()+
  theme_bw()

TukeyHSD(aov(ord.merged$X1 ~ ord.merged$Population.Code))

# Hausdorf & Hennig followed: 4 dimensions retained so no more than five individuals needed to recognize a cluster
head(wide.ldms)
wide.ldms1 <- wide.ldms[-(grep("TC", wide.ldms$Ind)),]

phdaisy_wide1 <- daisy(wide.ldms1[,-1], metric="gower")
phclust_wide1 <- hclust(phdaisy_wide1)

ph.dendwide1 <- as.dendrogram(phclust_wide1)

ord <- isoMDS(phdaisy_wide1)
code <- substr(wide.ldms1$Ind,1,2)

ord.xy1 <- data.frame(wide.ldms1$Ind,ord$points,code)
colnames(ord.xy1) <- c("Individual","X","Y","Site")

ggplot(ord.xy1, aes(X,Y,colour=Site))+
  geom_point()+
  stat_ellipse()+
  theme_bw()

#####


```

```{r}

#par(mar=c(5,5,2,2))
ord.1 <- isoMDS(ph.daisy)

ord.xy.1 <- data.frame(ph.merg,ord.1$points)

ggplot(ord.xy.1, aes(X1,X2,colour=Species))+
  geom_point()+
  stat_ellipse()+
  theme_bw()

###vegan vignette
ord.meta <- metaMDS(ph.daisy)

plot(ord.meta)

plot(ord,meta, type = "n")
points(ord.meta, display = "sites", cex = 0.8, pch=21, col="red", bg="yellow")
text(ord.meta, display = "spec", cex=0.7, col="blue")


#############
fit <- isoMDS(ph.daisy)
fit

plot(fit$points[,1], fit$points[,2], type="n")
text(fit$points[,1], fit$points[,2],labels = ph.merg$Population.Code, cex = 0.7)
```


Klingenberg et al 2011
```{r}


```


Landmark data assuming three species   
```{r}
# Phacelia formosula: BR, RN, and AS
# Phacelia scullyi: FR
# Phacelia gina-glennea: TC

wide.phfo <- reshape(rbind(landmarks[[1]],landmarks[[2]],landmarks[[4]]),
                     idvar="Ind", timevar = "Ldm", direction = "wide")
#wide.FC
#wide.TC

write.table(wide.phfo[,-1], "C:/Users/deprengm/Dropbox/Denver Botanic Gardens Postdoctoral Genetic Files/Phacelia project/Phacelia 2014 collection and morphology data/wide_phfo.txt",
            col.names=FALSE,row.names = FALSE, quote=FALSE)


new.dat.list2 <- lapply(c("wide_phfo","wide_FC","wide_TC"), function(x){
  read.in(paste("C:/Users/deprengm/Dropbox/Denver Botanic Gardens Postdoctoral Genetic Files/Phacelia project/Phacelia 2014 collection and morphology data/",x,".txt", sep=""),k,m)
})

# before procustes correction
plotshapes(new.dat.list2[[1]])
plotshapes(new.dat.list2[[2]])
plotshapes(new.dat.list2[[3]])

out.procf.species <- lapply(new.dat.list2, function(x){
  procGPA(x)
})

par(mfrow=c(1,3))
par(mar=c(2,2,2,2))
lapply(out.procf.species, function(x){
  plotshapes(x$rotated)
})

dev.off()

#Riemannian shape distance between mean shapes
species <- c("phfo","phsc","phgi")
compspecies <- t(combn(1:3,2))
reimannian.species <- lapply(1:3, function(x){
  r.s <- riemdist(out.procf.species[[ compspecies[x,][1]]]$mshape,
                  out.procf.species[[ compspecies[x,][2]]]$mshape)
  out <- cbind(paste(species[ compspecies[x,][1]],
                     species[ compspecies[x,][2]]),
               r.s)
})
r.s.species <- do.call(rbind, reimannian.species)
data.frame(r.s.species)


#test for 2D mean shape differences, boostrap and permutation tests
testshapediff.sp <- lapply(1:3, function(x){
  resampletest(new.dat.list2[[ compspecies[x,][1]]],
               new.dat.list2[[ compspecies[x,][2]]])
})

testshapediff.sp[[1]]

tsd.s <- lapply(1:3, function(x){
  out <- paste(species[ compspecies[x,][1] ],species[ compspecies[x,][2] ], sep="-")
  list(testshapediff.sp[[x]],out)
  })

tsd.s[[3]][[2]][1]

do.call(rbind,lapply(tsd.s, function(f){
  data.frame(SpComp = f[[2]][1],
             Lambda = round(f[[1]]$lambda,2),
             Lambda.pvlaue = round(f[[1]]$lambda.pvalue,2),
             James = round(f[[1]]$J,2),
             James.pvalue = round(f[[1]]$J.pvalue,2))
  })
)



```




  