---
title: "Zapata_Phacelia"
author: "Michelle DePrenger-Levin"
date: "November 29, 2016"
output: html_document
---

#Load libraries.
```{r}
library(labdsv)
#library(ellipse)
library(mvtnorm)
library(ggplot2)
```

```{r}
phfo <- read.csv("Q:/Research/Lab/Projects/Phacelia formosula/Morphology Images_phfo/Phacelia.formosula.data.2014CSV.csv")

head(phfo)
phfo <- phfo[,-(18:19)]
```

R script from Appendix 5 in Zapata F., Jimenez I. Species Delimitation in Systematics: Inferring Gaps in Morphology across Geography.
```{r}
# Principal component analysis on correlation matrix
# Getting rid of Species through Plant Height 
phfocc <- phfo[complete.cases(phfo),]
pa <- pca(phfocc[,-(1:9)], cor = TRUE)

# Add PC scores to the data matrix
phfocc <- cbind(phfocc, as.data.frame(pa$scores))

table(phfocc$Population.Code)
```

Calculate sample size, mean, and variance-covariance for each group
```{r}
dataphfo <- lapply(unique(phfocc$Population.Code), function(site){
  n_Site <- length(phfocc$PC1[phfocc$Population.Code == site])
  m1 <- array(c(mean(phfocc$PC1[phfocc$Population.Code == site]),
                mean(phfocc$PC2[phfocc$Population.Code == site])),
              dim = c(2,1))
  Z1 <- var(cbind(phfocc$PC1[phfocc$Population.Code == site], 
                  phfocc$PC2[phfocc$Population.Code == site]))
  data.frame(site, Comp = c("PC1","PC2"), n_Site, m1, Z1)
})

data.phfo <- do.call(rbind, dataphfo)
```

Ridgeline manifold (i.e., d<-a %*% b) from equation 4 in Ray and Lindsay (2005:2045) is evaluated at various values of alfa and the resulting coordinates (in PC1 and PC2 space) are captured in vectors x and y.  
Probability density function along the ridgeline manifold (pdf)

```{r}

df <- data.frame(matrix(combn(unique(data.phfo$site),2),2,10))
# df will have each population combination to compare

ridgemani <- apply(df, 2, function(site){
  
# apply over values of alpha  
  lst1 <- lapply(seq(0,1,0.001), function(alfa){
    Z1 <- data.phfo[data.phfo$site == as.character(site[1]),c("X1","X2")]
    Z2 <- data.phfo[data.phfo$site == as.character(site[2]),c("X1","X2")]
    m1 <- data.phfo[data.phfo$site == as.character(site[1]),"m1"]
    m2 <- data.phfo[data.phfo$site == as.character(site[2]),"m1"]
    a <- solve((1-alfa)*solve(Z1) + alfa*solve(Z2))
    b <- (1-alfa)*solve(Z1)%*%m1 + alfa*solve(Z2)%*%m2
    d <- a %*% b
    data.frame(site = paste(site[1], site[2], sep="-"),d[1], d[2])
    })
  
  s2s <- do.call(rbind, lst1)
  
                              # matrix of quantiles - each row
  s2s$pdf <- (1/2*(dmvnorm(cbind(s2s$d.1.,s2s$d.2.),  
                            #mean
                            data.phfo[data.phfo$site == as.character(site[1]),"m1"],
                            #covariance matrix
                            as.matrix(data.phfo[data.phfo$site == as.character(site[1]),c("X1","X2")])) +
                    dmvnorm(cbind(s2s$d.1.,s2s$d.2.),
                            data.phfo[data.phfo$site == as.character(site[2]),"m1"],
                            as.matrix(data.phfo[data.phfo$site == as.character(site[2]),c("X1","X2")]))))
  
  s2s <- data.frame(s2s, alfa = seq(0,1,0.001))
  })

ridge <- do.call(rbind, ridgemani)
head(ridge)
```


```{r}
phfo2 <- reshape(data.phfo, idvar = "site", timevar = "Comp", direction = "wide")

phfo2
```


Plot the first two components, the bivariate means of the two species/sites, and the ridgeline manifolds
```{r}
ggplot()+
  geom_point(data = phfocc, aes(PC1, PC2, colour = Population.Code))+
  stat_ellipse(data = phfocc, aes(PC1, PC2, colour = Population.Code))+
  geom_point(data = phfo2, aes(m1.PC1, m1.PC2, colour = site, cex = 3))

```

```{r}
ggplot(ridge, aes(alfa, pdf, colour = site))+
  geom_line()

```


Tolerance ellipses and corresponding beta values for various points along the ridgeline manifold.  
```{r}
TolEll <- apply(df, 2, function(site){
  p<-2
  #Change the value of gamma to the desired level of statistical confidence to calculate statistical tolerance regions (see Krishnamoorty and Mathew 1999)
  gamma<-0.95
  #Set variables per species
  n_SPA <- unique(data.phfo$n_Site[data.phfo$site == as.character(site[1])])
  n_SPB <- unique(data.phfo$n_Site[data.phfo$site == as.character(site[2])])
  x <- ridge$d.1.[ridge$site == paste(site[1], site[2], sep="-")]
  y <- ridge$d.2.[ridge$site == paste(site[1], site[2], sep="-")]
  m1 <- data.phfo$m1[data.phfo$site == as.character(site[1])]
  m2 <- data.phfo$m1[data.phfo$site == as.character(site[2])]
  Z1 <- data.phfo[data.phfo$site == as.character(site[1]),c("X1","X2")]
  Z2 <- data.phfo[data.phfo$site == as.character(site[2]),c("X1","X2")]
    
  k <- do.call(rbind,lapply(seq(2,1000,1), function(i){
    k1 <- (n_SPA-1)%*%t(c(x[i],y[i])-m1)%*%solve((n_SPA-1)*Z1)%*%(c(x[i],y[i])-m1)
    k2 <- (n_SPB-1)%*%t(c(x[i],y[i])-m2)%*%solve((n_SPB-1)*Z2)%*%(c(x[i],y[i])-m2)
    cbind(k1,k2)
  }))
  
    #the chi quantiles and beta values are obtained using the approximation in equation 3.15 in Krishnamoorty and Mathew (1999, page 239).
    chi_quantile<-k[,1]*(1/(n_SPA-1))*(qchisq(1-gamma, n_SPA-p))
    beta_or_chi_probability_SPA<-pchisq(chi_quantile, p, ncp=p/n_SPA, lower.tail = TRUE, log.p = FALSE)
    chi_quantile2<-k[,2]*(1/(n_SPB-1))*(qchisq(1-gamma, n_SPB-p))
    beta_or_chi_probability_SPB<-pchisq(chi_quantile2, p, ncp=p/n_SPB, lower.tail = TRUE, log.p = FALSE)
    #Gather in vectors the beta values for various points along the ridgeline manifold. 
    data.frame(site = paste(site[1],site[2],sep="-"), beta_vector_SPA = beta_or_chi_probability_SPA,
               beta_vector_SPB = beta_or_chi_probability_SPB,
               alfa = seq(0,1,0.001)[2:1000])
  })


```

```{r}
for(i in 1:10){
  tiff(paste("Q:/Research/Lab/Projects/Phacelia formosula/Morphology Images_phfo/",i,
            "betavalues.jpeg",sep=""), width = 84, height = 84, compression = "lzw", 
       units = "mm", res = 300)
    paste(ggplot(TolEll[[i]])+
            geom_point(aes(alfa, beta_vector_SPA), col = "red")+
            geom_point(aes(alfa, beta_vector_SPB), col = "purple")+
            ggtitle(unique(TolEll[[i]]$site))+
            theme_bw())
    dev.off()
}


```


```{r}
quantsa.b <- lapply(unique(phfocc$Population.Code), function(pop){
  phfo.pop <- phfocc[phfocc$Population.Code == pop,]
  quant <- lapply(8:25, function(variable){
    Qlow <- quantile(phfo.pop[,variable], 0.95)
    Qhigh <- quantile(phfo.pop[,variable], 0.05)
    data.frame(site = pop, 
               Variable = names(phfo.pop)[variable],
               Qlow, Qhigh)
    })
  do.call(rbind,quant)
  })

quant <- do.call(rbind, quantsa.b)


```

```{r}
ggplot()+
  geom_density(data = phfocc, aes(Plant.Height, fill = Population.Code),
               alpha = 0.25)+
  geom_vline(data = quant[quant$Variable == "Plant.Height",], 
             aes(xintercept = Qlow, colour = site), linetype = "dashed")+
  geom_vline(data = quant[quant$Variable == "Plant.Height",], 
             aes(xintercept = Qhigh, colour = site))

    
```

```{r}
ggplot()+
  geom_density(data = phfocc, aes(Width, fill = Population.Code),
               alpha = 0.25)+
  geom_vline(data = quant[quant$Variable == "Width",], 
             aes(xintercept = Qlow, colour = site), linetype = "dashed")+
  geom_vline(data = quant[quant$Variable == "Width",], 
             aes(xintercept = Qhigh, colour = site))

    
```


```{r}
ggplot()+
  geom_density(data = phfocc, aes(Pinnae.Pairs, fill = Population.Code),
               alpha = 0.25)+
  geom_vline(data = quant[quant$Variable == "Pinnae.Pairs",], 
             aes(xintercept = Qlow, colour = site), linetype = "dashed")+
  geom_vline(data = quant[quant$Variable == "Pinnae.Pairs",], 
             aes(xintercept = Qhigh, colour = site))

    
```

```{r}
ggplot(phfocc, aes(Pinnae.Pairs, fill = Species))+
  geom_density(alpha = 0.5)

```

```{r}
ggplot(phfocc, aes(PC1, fill = Species))+
  geom_density(alpha = 0.5)

```

```{r}
ggplot()+
  geom_density(data = phfocc, aes(PC1, fill = Population.Code),
               alpha = 0.25)+
  geom_vline(data = quant[quant$Variable == "PC1",], 
             aes(xintercept = Qlow, colour = site), linetype = "dashed")+
  geom_vline(data = quant[quant$Variable == "PC1",], 
             aes(xintercept = Qhigh, colour = site))

    
```



Grahic et al 2013 Morphological evalutation of common bean diversity in Bosnia and Herzegovina using he discriminant analysis of principal compenents (DAPC) multivariate method 


