---
title: "Zapata_Phacelia"
author: "Michelle DePrenger-Levin"
date: "November 29, 2016"
output: html_document
---

#Load libraries.
```{r}
library(labdsv)
library(ellipse)
library(mvtnorm)
library(ggplot2)
library(vegan)
library(spdep)
```

```{r}
phfo <- read.csv("Q:/Research/Lab/Projects/Phacelia formosula/Morphology Images_phfo/Phacelia.formosula.data.2014CSV.csv")

head(phfo)
phfo <- phfo[,-(18:19)]
```

R script from Appendix 5 in Zapata F., Jimenez I. Species Delimitation in Systematics: Inferring Gaps in Morphology across Geography.
```{r}
# Principal component analysis on correlation matrix
# Getting rid of Species through Plant Height 
phfocc <- phfo[complete.cases(phfo),]
pairs(phfocc[,-(1:7)])
cor(phfocc[,-(1:7)])

pa <- pca(phfocc[,-(1:7)], cor = TRUE)

summary(pa)
loadings(pa)

# Add PC scores to the data matrix
phfocc <- cbind(phfocc, as.data.frame(pa$scores))

table(phfocc$Population.Code)
```

Calculate sample size, mean, and variance-covariance for each group
```{r}
dataphfo <- lapply(unique(phfocc$Population.Code), function(site){
  n_Site <- length(phfocc$PC1[phfocc$Population.Code == site])
  m1 <- array(c(mean(phfocc$PC1[phfocc$Population.Code == site]),
                mean(phfocc$PC2[phfocc$Population.Code == site])),
              dim = c(2,1))
  Z1 <- var(cbind(phfocc$PC1[phfocc$Population.Code == site], 
                  phfocc$PC2[phfocc$Population.Code == site]))
  data.frame(site, Comp = c("PC1","PC2"), n_Site, m1, Z1)
})

data.phfo <- do.call(rbind, dataphfo)
```

Ridgeline manifold (i.e., d<-a %*% b) from equation 4 in Ray and Lindsay (2005:2045) is evaluated at various values of alfa and the resulting coordinates (in PC1 and PC2 space) are captured in vectors x and y.  
Probability density function along the ridgeline manifold (pdf)

```{r}

df <- data.frame(matrix(combn(unique(data.phfo$site),2),2,10))
# df will have each population combination to compare

ridgemani <- apply(df, 2, function(site){
  
# apply over values of alpha  
  lst1 <- lapply(seq(0,1,0.001), function(alfa){
    Z1 <- data.phfo[data.phfo$site == as.character(site[1]),c("X1","X2")]
    Z2 <- data.phfo[data.phfo$site == as.character(site[2]),c("X1","X2")]
    m1 <- data.phfo[data.phfo$site == as.character(site[1]),"m1"]
    m2 <- data.phfo[data.phfo$site == as.character(site[2]),"m1"]
    a <- solve((1-alfa)*solve(Z1) + alfa*solve(Z2))
    b <- (1-alfa)*solve(Z1)%*%m1 + alfa*solve(Z2)%*%m2
    d <- a %*% b
    data.frame(site = paste(site[1], site[2], sep="-"),d[1], d[2])
    })
  
  s2s <- do.call(rbind, lst1)
  
                              # matrix of quantiles - each row
  s2s$pdf <- (1/2*(dmvnorm(cbind(s2s$d.1.,s2s$d.2.),  
                            #mean
                            data.phfo[data.phfo$site == as.character(site[1]),"m1"],
                            #covariance matrix
                            as.matrix(data.phfo[data.phfo$site == as.character(site[1]),c("X1","X2")])) +
                    dmvnorm(cbind(s2s$d.1.,s2s$d.2.),
                            data.phfo[data.phfo$site == as.character(site[2]),"m1"],
                            as.matrix(data.phfo[data.phfo$site == as.character(site[2]),c("X1","X2")]))))
  
  s2s <- data.frame(s2s, alfa = seq(0,1,0.001))
  })

ridge <- do.call(rbind, ridgemani)
head(ridge)
```


```{r}
phfo2 <- reshape(data.phfo, idvar = "site", timevar = "Comp", direction = "wide")

phfo2
```


Plot the first two components, the bivariate means of the two species/sites, and the ridgeline manifolds
```{r}
ggplot()+
  geom_point(data = phfocc, aes(PC1, PC2, colour = Population.Code))+
  stat_ellipse(data = phfocc, aes(PC1, PC2, colour = Population.Code))+
  geom_point(data = phfo2, aes(m1.PC1, m1.PC2), col = "black", cex = 4.5)+
  geom_point(data = phfo2, aes(m1.PC1, m1.PC2, colour = site), cex = 3)+
  theme_bw()

```

```{r}

plot(phfocc$PC1, phfocc$PC2, cex=2, xlab="PC1", ylab="PC2", type = "n")
points(phfocc$PC1[phfocc$Population.Code=="FC"], 
       phfocc$PC2[phfocc$Population.Code=="FC"], cex=1, col="blue")
points(phfocc$PC1[phfocc$Population.Code=="BR"], 
       phfocc$PC2[phfocc$Population.Code=="BR"], cex=1, col="goldenrod")
points(data.phfo$m1[data.phfo$site == "FC"][1],
       data.phfo$m1[data.phfo$site == "FC"][2], cex=2, pch=19)
points(data.phfo$m1[data.phfo$site == "BR"][1],
       data.phfo$m1[data.phfo$site == "BR"][2], cex=2, pch=19)
points(ridge$d.1.[ridge$site == "FC-BR"], ridge$d.2.[ridge$site == "FC-BR"], 
       type="l", col="red", lty=1, lwd=3)

```

```{r}
ggplot(ridge, aes(alfa, pdf, colour = site))+
  geom_line()

```


Tolerance ellipses and corresponding beta values for various points along the ridgeline manifold.  
```{r}
TolEll <- apply(df, 2, function(site){
  p<-2
  #Change the value of gamma to the desired level of statistical confidence to calculate statistical tolerance regions (see Krishnamoorty and Mathew 1999)
  gamma<-0.95
  #Set variables per species
  n_SPA <- unique(data.phfo$n_Site[data.phfo$site == as.character(site[1])])
  n_SPB <- unique(data.phfo$n_Site[data.phfo$site == as.character(site[2])])
  x <- ridge$d.1.[ridge$site == paste(site[1], site[2], sep="-")]
  y <- ridge$d.2.[ridge$site == paste(site[1], site[2], sep="-")]
  m1 <- data.phfo$m1[data.phfo$site == as.character(site[1])]
  m2 <- data.phfo$m1[data.phfo$site == as.character(site[2])]
  Z1 <- data.phfo[data.phfo$site == as.character(site[1]),c("X1","X2")]
  Z2 <- data.phfo[data.phfo$site == as.character(site[2]),c("X1","X2")]
    
  k <- do.call(rbind,lapply(seq(2,1000,1), function(i){
    k1 <- (n_SPA-1)%*%t(c(x[i],y[i])-m1)%*%solve((n_SPA-1)*Z1)%*%(c(x[i],y[i])-m1)
    k2 <- (n_SPB-1)%*%t(c(x[i],y[i])-m2)%*%solve((n_SPB-1)*Z2)%*%(c(x[i],y[i])-m2)
    cbind(k1,k2)
  }))

    #the chi quantiles and beta values are obtained using the approximation in equation 3.15 in Krishnamoorty and Mathew (1999, page 239).
    chi_quantile<-k[,1]*(1/(n_SPA-1))*(qchisq(1-gamma, n_SPA-p))
    beta_or_chi_probability_SPA<-pchisq(chi_quantile, p, ncp=p/n_SPA, lower.tail = TRUE, log.p = FALSE)
    chi_quantile2<-k[,2]*(1/(n_SPB-1))*(qchisq(1-gamma, n_SPB-p))
    beta_or_chi_probability_SPB<-pchisq(chi_quantile2, p, ncp=p/n_SPB, lower.tail = TRUE, log.p = FALSE)
    #Gather in vectors the beta values for various points along the ridgeline manifold. 
    data.frame(site = paste(site[1],site[2],sep="-"),
               beta_vector_SPA = beta_or_chi_probability_SPA,
               beta_vector_SPB = beta_or_chi_probability_SPB,
               alfa = seq(0,1,0.001)[2:1000])
  })


toleranceell <- do.call(rbind, TolEll)

```

```{r}
ggplot(toleranceell, aes(colour = site))+
  geom_point(aes(alfa, beta_vector_SPA), col = "red")+
  geom_point(aes(alfa, beta_vector_SPB), col = "purple")+
  facet_grid(~site)+
  theme_bw()

```


```{r}
quantsa.b <- lapply(unique(phfocc$Population.Code), function(pop){
  phfo.pop <- phfocc[phfocc$Population.Code == pop,]
  quant <- lapply(8:25, function(variable){
    Qlow <- quantile(phfo.pop[,variable], 0.95)
    Qhigh <- quantile(phfo.pop[,variable], 0.05)
    data.frame(site = pop, 
               Variable = names(phfo.pop)[variable],
               Qlow, Qhigh)
    })
  do.call(rbind,quant)
  })

quant <- do.call(rbind, quantsa.b)


```

```{r}
ggplot()+
  geom_density(data = phfocc, aes(Plant.Height, fill = Population.Code),
               alpha = 0.25)+
  geom_vline(data = quant[quant$Variable == "Plant.Height",], 
             aes(xintercept = Qlow, colour = site), linetype = "dashed")+
  geom_vline(data = quant[quant$Variable == "Plant.Height",], 
             aes(xintercept = Qhigh, colour = site))+
  theme_bw()

    
```

```{r}
ggplot()+
  geom_density(data = phfocc, aes(Width, fill = Population.Code),
               alpha = 0.25)+
  geom_vline(data = quant[quant$Variable == "Width",], 
             aes(xintercept = Qlow, colour = site), linetype = "dashed")+
  geom_vline(data = quant[quant$Variable == "Width",], 
             aes(xintercept = Qhigh, colour = site))+
  theme_bw()

    
```


```{r}
ggplot()+
  geom_density(data = phfocc, aes(Pinnae.Pairs, fill = Population.Code),
               alpha = 0.25)+
  geom_vline(data = quant[quant$Variable == "Pinnae.Pairs",], 
             aes(xintercept = Qlow, colour = site), linetype = "dashed")+
  geom_vline(data = quant[quant$Variable == "Pinnae.Pairs",], 
             aes(xintercept = Qhigh, colour = site))

    
```

```{r}
ggplot(phfocc, aes(Pinnae.Pairs, fill = Species))+
  geom_density(alpha = 0.5)

```

```{r}
ggplot(phfocc, aes(PC1, fill = Species))+
  geom_density(alpha = 0.5)

```

Estimated phenotypic overlap between groups. 
```{r}
ggplot()+
  geom_density(data = phfocc, aes(PC1, fill = Population.Code),
               alpha = 0.25)+
  geom_vline(data = quant[quant$Variable == "PC1",], 
             aes(xintercept = Qlow, colour = site), linetype = "dashed")+
  geom_vline(data = quant[quant$Variable == "PC1",], 
             aes(xintercept = Qhigh, colour = site))

    
```



```{r}
data <- phfocc
#Assign codes to the groups to be compared
SPA <- "gina-gleneae"
SPB <- "formosula"
SPC <- "scullyi"

pops <- unique(phfocc$Population.Code)

#Assign colors to each species for visualization.
col_SPA<-c("red")
col_SPB<-c("blue")
col_SPC<-c("purple")

col_pops <- c("midnightblue","mediumvioletred","lightskyblue4","mediumseagreen","orangered")
```

function to compare and visualize two populations
```{r}
ridgeline.manifold <- function(pop1, pop2){
  n_SPA<-length(data$PC1[data$Population.Code==pops[pop1]])
m1 <- array(c(mean(data$PC1[data$Population.Code==pops[pop1]]),
              mean(data$PC2[data$Population.Code==pops[pop1]])),dim=c(2,1))
Z1<-var(cbind(data$PC1[data$Population.Code==pops[pop1]], data$PC2[data$Population.Code==pops[pop1]]))


#Calculate sample size, mean and variance-covariance matrix for SPB.
n_SPB<-length(data$PC1[data$Population.Code==pops[4]])
m2 <- array(c(mean(data$PC1[data$Population.Code==pops[pop2]]),
              mean(data$PC2[data$Population.Code==pops[pop2]])), dim=c(2,1))
Z2<-var(cbind(data$PC1[data$Population.Code==pops[pop2]], data$PC2[data$Population.Code==pops[pop2]]))

#Calculate ridgeline manifold. The equation for the manifold (i.e., d<-a %*% b) 
#   from equation 4 in Ray and Lindsay (2005:2045) is evaluated at various values 
#   of alfa and the resulting coordinates (in PC1 and PC2 space) are captured in 
#   vectors x and y.
x<-c()
y<-c()
alfa<-seq(0,1,0.001)        # 142 rows in phfo after removing NA indiviudals
for (i in 1:length(alfa)){
  a<-solve((1-alfa[i])*solve(Z1)  +  alfa[i]*solve(Z2))
  b<-(1-alfa[i])*solve(Z1)%*%m1  +  alfa[i]*solve(Z2)%*%m2
  d<-a %*% b
  x<-append(x, d[1], after=length(x))
  y<-append(y, d[2], after=length(y))
}

data1<-data[data$Population.Code %in% paste(pops[c(pop1,pop2)],sep=","),]


plot1 <- ggplot() +
  geom_point(data = data1, aes(PC1, PC2, colour = Population.Code)) +
  stat_ellipse(data = data1, aes(PC1, PC2, colour = Population.Code), 
               type = "norm") +
  stat_ellipse(data = data1, aes(PC1, PC2, colour = Population.Code), 
               level = 0.85, type = "norm") +
  stat_ellipse(data = data1, aes(PC1, PC2, colour = Population.Code), 
               level = 0.75, type = "norm") +
  stat_ellipse(data = data1, aes(PC1, PC2, colour = Population.Code), 
               level = 0.65, type = "norm") +
  geom_line(data = data.frame(x,y), aes(x,y), cex = 1.5) +
  theme_bw()

#beta values along ridgeline manifold  
  beta_vector_SPA<-c()
  beta_vector_SPB<-c()
  
  for (i in seq(2, 1000, 1)){
#calculation of ellipses and beta for SPA. The line below is equation 2.2 in Krishnamoorty and Mathew (1999, page 236).
    k<- (n_SPA-1)%*%t(c(x[i],y[i])-m1)%*%solve((n_SPA-1)*Z1)%*%(c(x[i],y[i])-m1)
    p<-2
    #Change the value of gamma to the desired level of statistical confidence to calculate statistical tolerance regions (see Krishnamoorty and Mathew 1999)
gamma<-0.95
#the chi quantiles and beta values are obtained using the approximation in equation 3.15 in Krishnamoorty and Mathew (1999, page 239).
chi_quantile<-k*(1/(n_SPA-1))*(qchisq(1-gamma, n_SPA-p))
beta_or_chi_probability_SPA<-pchisq(chi_quantile, p, ncp=p/n_SPA, lower.tail = TRUE, log.p = FALSE)

#calculation of ellipses and beta for SPB. The line below is equation 2.2 in Krishnamoorty and Mathew (1999, page 236).
k<- (n_SPB-1)%*%t(c(x[i],y[i])-m2)%*%solve((n_SPB-1)*Z2)%*%(c(x[i],y[i])-m2)
k
p<-2
#Change the value of gamma to the desired level of statistical confidence to calculate statistical tolerance regions (see Krishnamoorty and Mathew 1999)
gamma<-0.95
#the chi quantiles and beta values are obtained using the approximation in equation 3.15 in Krishnamoorty and Mathew (1999, page 239).
chi_quantile<-k*(1/(n_SPB-1))*(qchisq(1-gamma, n_SPB-p))
beta_or_chi_probability_SPB<-pchisq(chi_quantile, p, ncp=p/n_SPB, lower.tail = TRUE, log.p = FALSE)
#Gather in vectors the beta values for various points along the ridgeline manifold. 
beta_vector_SPA<-append(beta_vector_SPA, beta_or_chi_probability_SPA, after=length(beta_vector_SPA))
beta_vector_SPB<-append(beta_vector_SPB, beta_or_chi_probability_SPB, after=length(beta_vector_SPB))
}


# probability density function (pdf) along the ridgeline manifold
  x<-c()
  y<-c()
  alfa<-seq(0,1,0.001)		# 142 rows in phfo after removing NA indiviudals
  for (i in 1:length(alfa)){
    a<-solve((1-alfa[i])*solve(Z1)  +  alfa[i]*solve(Z2))
    b<-(1-alfa[i])*solve(Z1)%*%m1  +  alfa[i]*solve(Z2)%*%m2
    d<-a %*% b
    x<-append(x, d[1], after=length(x))
    y<-append(y, d[2], after=length(y))
    }
    
    par(mar=c(5,5,2,2))
  plot3 <- plot(alfa, (1/2*( dmvnorm(cbind(x,y) , m1, Z1) + dmvnorm(cbind(x,y) , m2, Z2) )), 
                type="l", ylab="Probability density", xlab=expression(~alpha), cex.lab=0.75,
                cex.axis=1, lwd=2)
  
  #Plot the value of beta at various points along the ridgeline manifold,
  #and an horizontal dashed line at beta=0.9 (=a possible a priori cutoff value of 0.1); change according to cutoff value defined a priori (see Wiens and Servedio 2000).
  plot2 <- recordPlot()

  plot(alfa[2:1000], beta_vector_SPA, type="l", col="black", ylim=c(0,1), lwd=2,
       xlab=expression(~alpha),
       ylab=expression(paste("Proportion ",beta, " within tolerance region")),
       cex.axis=1, cex.lab=1)
  points(alfa[2:1000], beta_vector_SPB, type="l", col="black", lty=1, lwd=2)
  points(alfa[1], beta_vector_SPA[1],cex=3, pch=22)
  text(alfa[1], beta_vector_SPA[1], pops[pop1])
  points(alfa[1], beta_vector_SPB[1],cex=3, pch=24)
  text(alfa[1], beta_vector_SPB[1], pops[pop2])
  points(alfa[999], beta_vector_SPA[999],cex=3, pch=22)
  points(alfa[999], beta_vector_SPB[999],cex=3, pch=24)
  abline(h=0.85, lty=3)   # abline(h=0.9, lty=3)

  list(plot1, plot2, plot3)
}

```

not working!! yargggg

maybe a do.call(tiff(...))

apply(df, 2, function(site){
tiff(paste("Q:/Research/Lab/Projects/Phacelia formosula/R code/R images/RidgelineManifold",
           site[1],"to",site[2],".tiff",sep=""),
width = 174, height = 150, compression="lzw", units = "mm", res=300)
  ridgeline.manifold(site[1],site[2])
  dev.off()
})
```{r}

ridgeline.manifold(1,2)[[1]]
ridgeline.manifold(1,3)[[3]]
ridgeline.manifold(1,4)[[2]]
ridgeline.manifold(1,5)[[3]]
ridgeline.manifold(2,3)[[3]]
ridgeline.manifold(2,4)[[2]]
ridgeline.manifold(2,5)[[1]]
ridgeline.manifold(3,4)[[3]]
ridgeline.manifold(3,5)[[3]]
ridgeline.manifold(4,5)[[3]]


par(mfrow=c(2,1))
ridgeline.manifold(1,2)[[1:2]]
ridgeline.manifold(1,3) 
ridgeline.manifold(1,4) 
ridgeline.manifold(1,5) 
ridgeline.manifold(2,3) 
ridgeline.manifold(2,4) 
ridgeline.manifold(2,5) 
ridgeline.manifold(3,4) 
ridgeline.manifold(3,5) 
ridgeline.manifold(4,5)

```

Geographic PCA   
1) look for statisctically signficant spatial linear trend and then   
2a) response variables are then the multivariate residuals around the linear trend    if not    
2b) response are original multivariate morphological measurements    
3) RDA - spatial eigenvectors are independent variables for hypothesis of geographic variation within a species - statistically signficant coefficient for the independent variables is support for two species hypothesis. 
```{r}
head(phfocc)

summary(pa)

```

Check for zero distance between points
```{r}

plot(phfocc$long,phfocc$lat,  cex=1.5)
n_coor<-nrow(phfocc)
factorial(n_coor)/(2*factorial(n_coor-2))
d<-as.matrix(dist(phfocc[,c("long","lat")]))
length(d[upper.tri(d)])
inspect_coor<-which(d==0, arr.ind=T)
nrow(inspect_coor)
(nrow(inspect_coor)-n_coor)/2 #25

inspect_coor

# move the points just a bit to get unique points per collection
sortphfo <- phfocc[ order(phfocc$long, phfocc$lat),]
sames <- rle(phfocc$long)
needtochange <- sames$values[sames$lengths > 1]

data.frame(from = phfocc$long[paste(needtochange, sep="|") %in% phfocc$long],
           to = jitter(phfocc$long[paste(needtochange, sep="|") %in% phfocc$long], 
                       amount = 0.00009))

phfocc$long[paste(needtochange, sep="|") %in% phfocc$long] <- jitter(phfocc$long[paste(needtochange, sep="|") %in% phfocc$long], 
                       amount = 0.00009)


data <- phfocc
```

```{r}

RDA <- function(pop1, pop2){
  #Run first RDA to detrend original response variables (PC1, PC2) using geographic coordinates as explanatory variables.
  #Define response variables.
  data_PC1<-data$PC1[data$Population.Code==pops[pop1] | 
                       data$Population.Code==pops[pop2]]
  data_PC2<-data$PC2[data$Population.Code==pops[pop1] | 
                       data$Population.Code==pops[pop2]]
  #Define explanatory variables.
  X<-data$long[data$Population.Code==pops[pop1] | 
                       data$Population.Code==pops[pop2]]
  Y<-data$lat[data$Population.Code==pops[pop1] | 
                       data$Population.Code==pops[pop2]]
  coor <- cbind(X,Y)
  n_coor<-nrow(phfocc[data$Population.Code==pops[pop1] | 
                       data$Population.Code==pops[pop2],])
  
  d <- as.matrix(dist(phfocc[data$Population.Code==pops[pop1] | 
                       data$Population.Code==pops[pop2],
                       c("long","lat")]))
  
  #Define the formula for RDA.
  b<-c("X","Y")
  formula_det<-as.formula(paste("cbind(data_PC1,data_PC2) ~ ", 
                                paste(b, collapse= "+")))
  #Run and plot RDA and test significance by permutation.
  rda_det_model<-rda(formula_det)
  plot(rda_det_model)
  anovaresult <- anova(rda_det_model, alpha=0.05, 
                       beta=0.01, step=100, perm.max=9999) #significant!
  
  #If (and only if) there is a siginificant linear trend, run the next script line
  #to extract residuals.
  ifanova <- if(anovaresult$`Pr(>F)`[1] < 0.05) {
    det_residuals<-residuals(rda_det_model)
    det_residuals
  }
  
  mst <- spantree(d)
  plot(coor)
  lines(mst, data.frame(X,Y))
  ws<-(d<=max(mst$dist))*(1-((d/(4*(max(mst$dist))))^2))
  w<-(d>0)*ws
  Id<-diag(n_coor)
  ones<-rep(1, n_coor)
  dim(ones)<-c(n_coor, 1)
  l<-ones%*%t(ones)
  l_n<-l/n_coor
  res1<-Id-l_n
  res2<-res1%*%w
  res3<-res2%*%res1
  ei_sim<-eigen(res3)
  
  #Run second RDA using eigenvectors associated with positive eigenvalues,
  #dummy variables for each species, and their interactions as explanatory variables.
  #Use either original variables or multivariate residuals as response variables, depending
  #on wether the first RDA yielded a siginificant linear trend.
  #List and enumerate eigenvectors associated with positive eigenvalues.
  pos_eivec<-seq(1:length(ei_sim$values))[ei_sim$values>0]
  n_pos_eivec<-length(ei_sim$values[ei_sim$values>0])
  #Define dummy variables and their interactions.
  dum <- as.numeric(data$Population.Code[data$Population.Code==pops[pop1] | 
                       data$Population.Code==pops[pop2]]==pops[pop1])
  interactions<-as.matrix(ei_sim$vectors[,1:n_pos_eivec]*dum)
  #Define the formula for RDA.
  a<-c(paste("ei_sim$vectors[,", pos_eivec, "]", sep=""), "dum",
       paste("interactions[,", c(1:n_pos_eivec), "]", sep=""))
  
  #If there is no significant linear trend in first RDA, then run next script line to use original PCA axes as response variables.
  formula1<-as.formula(paste("cbind(data_PC1,data_PC2) ~ ", 
                             paste(a, collapse= "+")))
  
  #If there was a significant linear trend in  first RDA, then run the next script line to use residuals as response variables.
  formula1<-as.formula(paste("det_residuals ~ ", paste(a, collapse= "+")))
  
  #Run and plot RDA, and test significance of different terms by permutation.
  rda_model1<-rda(formula1)
  plot(rda_model1)
  rda1 <- anova(rda_model1, alpha=0.05, beta=0.01, step=100, perm.max=9999, by = "terms")
  
  #The lines below apply standard model simplification procedures (Crawley 2002).
  rda_model2<-update(rda_model1, ~.-interactions[, 1])
  anova(rda_model2, alpha=0.05, beta=0.01, step=100, perm.max=9999, by = "terms")
  
  #The lines below can be used to obtain adjusted redundancy statistic (Peres-Neto et al. 2006)
  d<-c(paste("ei_sim$vectors[,", pos_eivec, "]", sep=""))
  e<-c("dum", paste("interactions[,", c(1:n_pos_eivec), "]", sep=""))
  
  #If there is no significant linear trend in first RDA, then run next script line to use original PCA axes as response variables.

  if(rda1$`Pr(>F)`[3] >= 0.05){
    formula2<-as.formula(paste("cbind(data_PC1,data_PC2) ~ ", 
                               paste(d, collapse= "+")))
    formula3<-as.formula(paste("cbind(data_PC1,data_PC2) ~ ", 
                               paste(e, collapse= "+")))
  }
    
    #If there was a significant linear trend in  first RDA, then run the next script line to use residuals as response variables.
  
  if(rda1$`Pr(>F)`[3] < 0.05){
    formula2<-as.formula(paste("det_residuals ~ ", paste(d, collapse= "+")))
    formula3<-as.formula(paste("det_residuals ~ ", paste(e, collapse= "+")))
  }
    
    #Update formula 2 or 3 according to simplification of rda_model
  formula3a<-update(formula3, ~.-interactions[,1])
  formula3a

  #calculate adjusted R-squared.
  variance_explained<-varpart(det_residuals, formula2, formula3)
  variance_explained

  
  
  list(Sites = cbind(pops[pop1],pops[pop2]),
       RDAPlot = rda_det_model, 
       anova = anovaresult, 
       residuals = ifanova,
       var_ex = variance_explained)
}


```

```{r}
numdf <- combn(1:5,2)

rdas <- apply(numdf, 2, function(site){
  RDA(site[1],site[2])
})

plot(rdas[[1]][[2]])

```
Grahic et al 2013 Morphological evalutation of common bean diversity in Bosnia and Herzegovina using he discriminant analysis of principal compenents (DAPC) multivariate method 


